---
title: "lab 5"
author: "Mari Herbst"
date: "10/26/2021"
output: 
  html_document: 
    toc: yes
    theme: spacelab
    number_sections: yes
    code_folding: hide
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Confidence intervals

In this section, we will learn to use the 't.test()' function as a way to return the confidence interval using the t-distribution.

## Example 1

Create a mock sample of observation values and find the confidence interval. Then, we'll learn how to do report that.

```{r}
pika_mass <- c(82, 76, 94, 102, 70, 86, 92, 110)

pika_mean <- mean(pika_mass)

pika_sd <- sd(pika_mass)

pika_ci <- t.test(pika_mass)

pika_ci
```

### Risky and tedious way to report

Mean pika mass at Paiute Pass is 89 g $\pm$ 13.27 g (mean $\pm$ 1 standard deviation, n = 7) with a 95% confidence interval of [77.9, 100.1]

### High initial investment but safer and better way

In-line code referencing to stored objects

Use a single backtick on either side of a lowercase r to create an in-line bit of code

Mean pika mass at Paiute Pass is `r round(pika_mean, 1)` g $\pm$ `r round(pika_sd, 1)` g (mean $\pm$ 1 standard deviation, n = `r length(pika_mass)`) with a 95% confidence interval of [`r round(pika_ci$conf.int[1], 1)`, `r round(pika_ci$conf.int[2], 1)`] g.

# Two-sample t-test

Use a two-sample t-test to test the null hypothesis that samples were drawn from populations with the same mean (difference in means = 0).

- H0: means difference = 0
- HA: means difference is NOT = 0

This is a two-sided t-test because no directionality is implied.

## Example 1

```{r}
oak_trees <- c(29, 37, 16, 22, 8, 18, 35, 24)

pine_trees <- c(48, 62, 69, 34, 30, 76, 19, 55)
```

Is there a significant difference in mean heights for oak and pine trees?

```{r}
trees_t <- t.test(oak_trees, pine_trees)
```

p-value `r trees_t$p.value` means that there is a `r (trees_t$p.value) * 100`% chance of finding sample means at least as different as those I found by random chance if they were drawn from populations with the same mean height.

